\chapter{AADL Properties}
\label{chap::aadl_properties}

This section gather basic function to manipulate the properties  defined in the property set First, we define those properties using plain AADL concepts, then we map them to a more precise set of Coq types. This two-step approach allows for a step-by-step refinement from an AADL model to a deep embedding in Coq terms.

\paragraph{Step 1: from AADL to formalized AADL in Coq} we first provide a definition of some AADL properties using the formalizations of AADL concepts from [aadl]. For each property, we define
\begin{itemize}
\item its name as an identifier [<property\_nane>\_Name]
\item the corresponding property type definition as [<property\_nane>]
\item a helper function [Is\_<property\_nane>] that returns true iff. a property\_value has property\_name as name.
\end{itemize}

Note: there is no need to check the type is valid, we assume the model is well-formed.

\paragraph{Step 2: deep Coq embedding} in the following, we provide concrete Coq type definitions, and a mapping from AADL concepts. This deeper embedding of AADL as Coq type will allow for more precise processing. For each property, we define

\begin{itemize}
\item its corresponding Coq type,
\item a function mapping the AADL property value to this type.
\end{itemize}

Note: these functions assumes the input AADL model elements are well-formed. In particular that property values are unique in a component.

\input{generated-content/aadl_communication_properties}
\input{generated-content/aadl_thread_properties}
